Abstract Data Type Inference

### 2015-03-22 formula.tex
More motivation paragraphs.

TODO MOTIVATE THIS SECTION

>>>>

While Section~\ref{sec:patterns} demonstrates that (the complements) of
naturally-occurring ADTs can be represented finitely, in this section we
demonstrate that such finite representations have logical interpretations,
allowing us to derive the formulas representing ADTs. In what follows, we
describe how to obtain a formula which is satisfied by the histories which
embed a given history, via $\preceq$. Then, using the finite set of histories
which represent the complement of a given ADT, we represent the ADT itself as
the conjunction of negations of these embedding formulas. The resulting formula
is satisfied only on the histories which do not embed the generators of a given
ADT’s complement.

### 2015-03-22 matching.tex
More motivation paragraphs.

TODO MOTIVATE THIS SECTION, AND ALSO ACCOUNT FOR THE READ-ONLY OPERATION
LABELING TOO.

A \emph{language} $L = \tup{E,P,Q}$ is a set $E$ of executions, along with
finite sequences $P = P_1 P_2 \ldots$ and $Q = Q_1 Q_2 \ldots$ of binary and
ternary predicates $P_i(e,o_1)$ and $Q_i(e,o_1,o_2)$ ranging over the
executions of $E$ and their operations. When $P$ is a $k$-length sequence of
$n$-ary predicates, we write $P(x_1, \ldots, x_n)$ to denote the valuation
sequence
\begin{align*}
  P_1(x_1, \ldots, x_n) \ldots P_k(x_1, \ldots, x_n).
\end{align*}
Given a language $L = \tup{E,P,Q}$, we say a matching scheme $M$ is
\emph{simple} when there exists an $n$-ary Boolean function $G$, for $n =
2\cdot|P|+|Q|$, such that for each execution $e \in E$
\begin{itemize}

@@

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  $L_\mathrm{reg}$
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{w}(e,o) & \text{$o$ is a {\sf write} operation in $e$, and}, \\
    & \mathsf{v}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  We define the function $G(x_\mathsf{w}, y_\mathsf{w}, z_\mathsf{v})$ over
  valuations to the predicates above to be satisfied if and only if:
  \begin{align*}
    x_\mathsf{w} \land z_\mathsf{v}
  \end{align*}
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined.

@@

  TODO SHOW THAT PREVIOUS SIMPLE MATCHING SCHEME NORMALIZES SOMETHING

>>>>

Our solution to the ADT inference problem relies on identifying faithful
matching schemes for ADT implementations. Though Section~\ref{sec:nature} shows
such matching schemes exist for naturally-occurring ADTs, the next natural
question with regard to automation is whether these matching schemes can be
generated automatically.

In this section we demonstrate that matching schemes can in fact be generated
systematically with minimal manual specification by reduction to logical
satisfiability. Essentially, we formulate matching schemes as uninterpreted
functions in satisfiability queries constrained by the requirement that they
\emph{normalize} an implementation’s executions, i.e.,~generate histories which
are closed under the operation removals of Section~\ref{sec:patterns}. This
requirement is given with respect to an enumeration of execution pairs in which
the first is admitted by the given implementation, and the second is not, yet
it is a projection of the first’s operations. Consequently, this prohibits any
normalizing matching scheme from considering the projected operations a match.
In what follows, we suppose the read-only component $R$ of matching schemes
$\tup{M,R}$ is given, and focus on the generation of the per-execution matching
functions $M$.

The required manual specification includes identifying a class of executions to
which an implementation should be subjected, and a set of predicates that are
required in the logic of matching schemes. Formally, a \emph{language} $L =
\tup{E,P,Q}$ is a set $E$ of executions, along with finite sequences $P = P_1
P_2 \ldots$ and $Q = Q_1 Q_2 \ldots$ of binary and ternary predicates
$P_i(e,o_1)$ and $Q_i(e,o_1,o_2)$ ranging over the executions of $E$ and their
operations. When $P$ is a $k$-length sequence of $n$-ary predicates, we write
$P(x_1, \ldots, x_n)$ to denote the valuation sequence
\begin{align*}
  P_1(x_1, \ldots, x_n) \ldots P_k(x_1, \ldots, x_n).
\end{align*}
Given a language $L = \tup{E,P,Q}$, we say a matching scheme $M$ is
\emph{simple} when there exists an $n$-ary Boolean function $G$, for $n =
2\cdot|P|+|Q|$, such that for each execution $e \in E$
\begin{itemize}

@@

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{w}(e,o) & \text{$o$ is a {\sf write} operation in $e$, and}, \\
    & \mathsf{v}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  We define the function $G(x_\mathsf{w}, y_\mathsf{w}, z_\mathsf{v})$ over
  valuations to the predicates above to be satisfied if and only if:
  \begin{align*}
    x_\mathsf{w} \land z_\mathsf{v}
  \end{align*}
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined. This
  scheme is exactly that which is specified in Example~\ref{ex:matching} of
  Section~\ref{sec:histories}.

@@
